\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper,scale=0.75}
\linespread{1.5}
\usepackage{graphicx} 
\usepackage{float} 
\usepackage{subfig} 
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{array}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{amsfonts}
\usepackage[english]{babel}
\usepackage{amsthm}
\usepackage{dcolumn}
\usepackage{multicol}
\usepackage{stfloats}
\usepackage{lscape}
\usepackage[figuresright]{rotating}
\RequirePackage{pdflscape}
\usepackage[toc,page]{appendix}
\usepackage{geometry}
\usepackage{longtable}
\usepackage{comment}
\usepackage{xcolor}

% -------- enumerated sub-labels (a), (b), … --
\usepackage{enumitem}
\setlist[enumerate,1]{label=(\alph*),ref=\alph*}
% ---------------------------------------------

\usepackage{hyperref}
\hypersetup{hidelinks,
	colorlinks=true,
	allcolors=black,
	pdfstartview=Fit,
	breaklinks=true}
\usepackage{csquotes}
\usepackage{natbib}
\bibliographystyle{apalike}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem*{remark}{Remark}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}
\newtheorem{assumption}{Assumption}[section] % number within sections


\begin{document}

\thispagestyle{empty}

\begin{center}
    THE HONG KONG UNIVERSITY OF SCIENCE AND TECHNOLOGY\\
    {\large \textbf{ECON 3123 Midterm Exam}}\\
    Date: Oct 10, 2025\\
    Time allowed: 80 minutes
\end{center}

\vspace{12pt}
\paragraph{Not to be taken away.}
\paragraph{Instructions:}
\begin{itemize}
    \item Answer ALL the questions. Write your answers on the answer book. Anything written on the question book will NOT be graded.
    \item Write your answer to all the multiple choice questions on one page together. Use one separate page for each short-answer question, and separate subquestions within one question by 1 to 2 blank lines. 
    \item Make sure that all your handweritngs are legible. Anything that cannot be understood by the grader will not be graded.
    \item Please submit BOTH the question book and the answer book after the exam.
\end{itemize}

\vspace{36pt}
\begin{center}
    \Large
    \textbf{DO NOT OPEN UNTIL INSTRUCTED!}
\end{center}

\vspace{36pt}
\subsection*{Name:}
\subsection*{Student ID: }

\clearpage
\setcounter{page}{1}
\subsection*{Multiple Choice Questions (20 points)}
4 points each. Choose the one alternative that best answers the question.
\begin{enumerate}[label=\arabic*.]
    \item  Suppose a country has 10 million people, of whom 8 million are working age. Of these 8 million, 5 million have jobs. Of the remainder: 1 million are actively searching for jobs; 1 million would like jobs but are not searching; and 1 million do not want jobs at all. The official unemployment rate is
    \begin{enumerate}[label=\Alph*.]
        \item 0.1
        \item 0.125
        \item 0.143
        \item 0.167
        \item 0.28
    \end{enumerate}
    \item A firm's value added equals
    \begin{enumerate}[label=\Alph*.]
        \item its revenue minus all of its costs.
        \item its revenue minus its wages.
        \item its revenue minus its wages and profit.
        \item its revenue minus its cost of intermediate goods.
        \item none of the above
    \end{enumerate}
    \item Suppose all the nominal variables are the same, if the expected inflation rate rises, the real PDV will
    \begin{enumerate}[label=\Alph*.]
        \item fall.
        \item rise.
        \item remain unchanged.
        \item none of the above
    \end{enumerate}
    \item Suppose that the recent fall of Silicon Valley Bank makes bank managers in all the other banks want to hold a higher fraction of deposit as reserve. This will cause:
    \begin{enumerate}[label=\Alph*.]
        \item an increase in the monetary base ($H$)
        \item a reduction in $H$ and a reduction in the money multiplier
        \item an increase in the money multiplier
        \item a reduction in the money multiplier
    \end{enumerate}
    \item If the central bank wants to cut the federal funds rate, it can:
    \begin{enumerate}[label=\Alph*.]
        \item increase the reserve requirement for banks.
        \item sell bonds in the market for bonds.
        \item buy bonds in the market for bonds. 
        \item increase the discount rate.
        \item none of the above.
    \end{enumerate}
\end{enumerate}

\newpage
\subsection*{Short-Answer Questions (80 points)}
Wrtie your calculation and explanation to get full scores. Partially correct steps earn partial credits.


\textbf{Question 6: Goods Market and Fiscal Policy (20 points)}

\textit{Consider an economy with the following behavior equations:}
\begin{align*}
    C &= c_0 + c_1Y_D\\
    I &= b_0 + b_1Y- b_2i\\
    T &= t_0 + t_1Y\\
    Y_D &= Y- T.
\end{align*}
\textit{$i$ is the nominal interest rate set by the central bank and $G$ is government spending.}

\textit{Assume that $t_1$ is between 0 and 1, and $c_1+b_1<0.5$.}

\begin{enumerate}[label=\alph*.]
    \item (5 points) Solve for equilibrium output and taxes.
\end{enumerate}

\textit{Suppose that the government starts with a balanced budget ($G=T$) and that there is a drop in $b_0$.}

\begin{enumerate}[label=\alph*., resume]
    \item (10 points) How should the government adjust its spending $G$ in order to keep the budget balanced? Does this government spending adjustment counteract or reinforce the effect of the drop in $b_0$ on output? \textbf{Explain your answers in words}.
    \item (5 points) After the adjustment of government spending $G$ that rebalances the budget, what happens to the equilibrium private saving?
\end{enumerate}

\newpage
\textbf{Question 7: Bank Crisis and Monetary Policy (30 points)}

\textit{Suppose that an economy is described by the following equations}
\begin{align*}
    C &= 0.5 + 0.2(Y-T) \\
    I &= 0.2 + 0.3Y -2.5(i+x) \\
    Z &= C + I + G\\
    G &= 1,\,\,\,\,\, T = 1\\
    i &= \bar{i}\\
    x &= 5\%\\
    \frac{M^d}{P} &= Y \times (0.7 - 4i)
\end{align*}
\textit{The current GDP deflator is 2. The public holds 20\% of the money in currency and the rest in checkable deposits. The average reserve ratio of the commercial banks is 25\%. }
\begin{enumerate}[label=\alph*.]
    \item (10 points) Derive the equilibrium output as a function of the nominal interest rate ($IS$ relation).

	\item (10 points) Suppose the nominal interest rate target $\bar{i}$ is 5\%. Compute the monetary base that can achieve the interest rate target. 
\end{enumerate}

\textit{Suppose there is a banking crisis which makes the average reserve ratio of the commercial banks increase from 25\% to 30\%. }
\begin{enumerate}[label=\alph*., resume]
    \item (5 points) Suppose that the central bank wants to restore the equilibrium nominal interest rate back to its target 5\%, how much should the new monetary base be?
\end{enumerate}

\textit{Suppose due to the same banking crisis, the risk premium x also increases from 5\% to 15\%. }
\begin{enumerate}[label=\alph*., resume]
    \item (5 points) Calculate the new monetary base that is necessary for the central bank to keep the nominal interest rate unchanged from its 5\% target.
\end{enumerate}

\newpage
\textbf{Question 8: Bond price and yield (30 points)}

\textit{Consider two types of bonds: a 2-year bond and a 3-year bond.  Both types have face value equal to \$100. }

\textit{The current risk free one-year interest rate is 4\% and the risk premium is 5\%. }

\textit{The market expects the central bank to cut risk-free one-year interest rate to 3\% next year and 2\% the year after.}

\begin{enumerate}[label=\alph*.]
    \item (10 points) What should be the current price for the 2-year bond? 
    \item (10 points) What should be the current price for the 3-year bond? 
    \item (10 points) What is the current yield to the 3-year bond?
\end{enumerate}

\vspace{300pt}
\begin{center}
    \textbf{\Large ******** END OF THE EXAM ********}
\end{center}

\end{document}